---
title: "p8105_hw6_zh2596"
author: "Zilin Huang"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(ggplot2)
library(modelr)
library(mgcv)
```

## Problem 1

Import and clean the homicide data

```{r}
homicide <- read.csv("data/homicide-data.csv")

homicide = 
  homicide |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1),
    victim_age = as.numeric(victim_age)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Fit a `glm` model for the city of Baltimore, MD

```{r}
homicide_glm = 
  filter(homicide, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _) |>
  broom::tidy() |>
  mutate(
    Odds_Ratio = exp(estimate), 
    Odds_Ratio_upper = exp(estimate + 1.96 * std.error),
    Odds_Ratio_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(Odds_Ratio, Odds_Ratio_lower, Odds_Ratio_upper)
```

Fit `glm` models for all cities

```{r}
model_cities = 
  homicide |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    Odds_Ratio = exp(estimate), 
    Odds_Ratio_upper = exp(estimate + 1.96 * std.error),
    Odds_Ratio_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, Odds_Ratio, Odds_Ratio_lower, Odds_Ratio_upper)
```

Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot

```{r}
model_cities |> 
  mutate(city_state = fct_reorder(city_state, Odds_Ratio)) |> 
  ggplot(aes(x = city_state, y = Odds_Ratio)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = Odds_Ratio_lower, ymax = Odds_Ratio_upper))
```

## Problem 2

Import the data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

```{r}
set.seed(1)
         
weather_df_strap =
  weather_df |>
  modelr::bootstrap(n = 5000) |>
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results_for_r_square = map(models, broom::glance),
    results_for_beta = map(models, broom::tidy)
  )

```

Generate estimates of the 5000 $r^2$ values, find its $2.5\%$ and $97.5\%$ quantiles, and plot its distribution

```{r}
weather_df_r_squared =
  weather_df_strap |>
  select(results_for_r_square) |> 
  unnest(results_for_r_square) |>
  select(r.squared) 

weather_df_r_squared |>      
  summarize(
    ci_lower = quantile(r.squared, 0.025), 
    ci_upper = quantile(r.squared, 0.975))

weather_df_r_squared |>
  ggplot(aes(x=r.squared)) + geom_histogram()
```

The distribution of $r^2$ is approximately symmetric around the value of 0.91.

Generate estimates of the 5000 $log(\beta_1 * \beta_2)$ values, find its $2.5\%$ and $97.5\%$ quantiles, and plot its distribution

```{r}
weather_df_beta =
  weather_df_strap |>
  select(results_for_beta) |>
  unnest(results_for_beta) |>
  select(-c(std.error, statistic, p.value)) |>
  filter(term != "(Intercept)") |>
  mutate(boot_id = rep(1:5000, each=2))  |>
  pivot_wider(
    names_from = "term",
    values_from = "estimate"
  ) |>
  mutate(
    log_beta1_beta_2 = log(tmin * prcp)
  ) |>
  select(log_beta1_beta_2)

weather_df_beta |>      
  summarize(
    ci_lower = quantile(log_beta1_beta_2, 0.025, na.rm = TRUE), 
    ci_upper = quantile(log_beta1_beta_2, 0.975, na.rm = TRUE))

weather_df_beta |>
  ggplot(aes(x=log_beta1_beta_2)) + geom_histogram()
```

The distribution of $log(\beta_1 * \beta_2)$ is approximately left-skewed.

## Problem 3

Import and clean the dataset

```{r}
birthweight = read.csv("data/birthweight.csv") |>
  drop_na() 

birthweight$babysex = as.factor(birthweight$babysex)
birthweight$malform = as.factor(birthweight$malform)
birthweight$frace = as.factor(birthweight$frace)
birthweight$mrace = as.factor(birthweight$mrace)
```

Based on the hypothesized structure for the factors that underly birthweight, I propose that babies and mothers' body status (like height, age at delivery, etc.) as well as their  smoking status may function as the coefficients affecting babies' birthweight. Thus, I make the following regression model.

```{r}
fit_1 = lm(bwt ~ bhead + blength + smoken + mheight + momage + ppbmi + ppwt , data = birthweight)
summary(fit_1)
```

I then show a plot of model residuals against fitted values using `add_predictions` and `add_residuals`

```{r}
birthweight |>
  add_predictions(fit_1) |>
  ggplot(aes(x = bwt, y = pred)) + geom_point()

birthweight |>
  add_residuals(fit_1) |>
  ggplot(aes(x = bwt, y = resid)) + geom_point()
```
Since the plot of real birthweight value against its predicted counterpart is linear, and the plot of real value against residuals is non-linear, I would deduce that this model fits well to predict babies' birthweight value.

Compare to other two models, the first one using length at birth and gestational age as predictors (main effects only), the second one using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
fit_2 = lm(bwt ~ blength + gaweeks, data = birthweight)
fit_3 = lm(bwt ~ bhead + blength + babysex + bhead*blength + 
           bhead*babysex + blength*babysex + blength*babysex*blength, data = birthweight)
```

Compare the three models using `crossv_mc` for generating cross-validated prediction errors

```{r}
cv_df = 
  crossv_mc(birthweight, 100) |>
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) |>
  mutate(
    mod_1 = map(train, \(df) fit_1),
    mod_2 = map(train, \(df) fit_2),
    mod_3 = map(train, \(df) fit_3)
  ) |> 
  mutate(
    rmse_1 = map2_dbl(mod_1, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_2 = map2_dbl(mod_2, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_3 = map2_dbl(mod_3, test, \(mod, df) rmse(model = mod, data = df)))

cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

It seems like model 2 contains the highest cross-validated prediction error value, implying that a model only considering one of babies' body status index (length at birth) and mothers' gestational age may not predict the birth weight well.